---
title: "The Tweets of Former President Obama"
---

```{r, warning=FALSE, message=FALSE, error=FALSE}
library(tidyverse)
hashtags <- read.csv("tweets.csv") |>
  mutate(text = str_extract_all(text, "#\\w+")) |>
  unnest(text) |>                         
  count(text, name = "num") |>                    
  rename(Hashtag = text) |>                    
  arrange(desc(num)) |>
  slice(1:7) |>
  mutate(Hashtag = fct_reorder(Hashtag, -num))

hashtags |>
  top_n(7, num) |>
  ggplot(aes(x = Hashtag, y = num)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(
    title = "Obama's Favorite Hashtags on Twitter",
    y = "Tweets",
    x = ""
  )
  
    
```
This is a great way to see what Obama's presidency was about. His most notable policy reforms are on this list. Over 600 tweets about #ActOnClimate—he really cared!

```{r, warning=FALSE, message=FALSE, error=FALSE}
read.csv("tweets.csv") |>
  mutate(wordCount = str_count(text, "\\S+")) |>
  ggplot(aes(x = wordCount)) +
  geom_histogram(binwidth = 1, fill = "white") +
  labs(
    title = "Word Count of Obama's Tweets",
    x = "Word Count",
    y = ""
  )
```

Obama's tweets have a beautiful distribution of word length. The Central Limit Theorem in action! I will say it is negatively skewed—but good enough for me. 18-19 words seems to be the sweet spot for Mr. Obama. Short and sweet.

```{r, warning=FALSE, message=FALSE, error=FALSE}
read.csv("tweets.csv") |>
  mutate(hour = str_extract(timestamp, "(?<=\\d{4}-\\d{2}-\\d{2} )\\d{2}")) |>
  mutate(hour = ifelse(as.numeric(hour) - 5 < 0, as.numeric(hour) + 19, as.numeric(hour) - 5)) |> 
  
  ggplot(aes(x = hour)) +
  geom_histogram(binwidth = 1, fill = "red") +
  labs(
    title = "The Time of Day Obama Likes to tweet",
    y = "",
    x = "Time EST"
  )
```

Obama tweets the most around mid-day, then picks it back up for a big spike at around 8pm. That must be when he gets the creative juices flowing! Would be interesting to research what makes 8pm such a popular tweeting time for him.

Data was collected by the Barack Obama Presidential Library to preserve his online presence: https://www.obamalibrary.gov/digital-research-room/archived-white-house-websites-and-social-media
