---
title: "Drunk Passenger"
---

There are n passengers getting on a fully booked plane. The first guy to board the plane got taken away at the terminal bar earlier and ended up losing his ticket. He walks on and randomly chooses a seat to sit in. Everyone else begins to board, one by one, if their assigned seat is avaible, they will sit there, if it has been taken, they do as the first guy and choose randomly. Whats the probability that the last guy to board sits in his assigned seat?

```{r, include=FALSE}
library(tidyverse)
library(purrr)

```

Let's simulate for planes ranging from 2 to 100 passengers.

```{r}
drunkPassenger <- function(n, seats){
  #generate random seat for first (drunk) passenger
  if (n == 1) {
    seats <-seats[seats != sample(seats, 1, replace = FALSE)]
  }
  
  #check if last seat matches last person, return true/false
  if (length(seats) == 1){
    return(seats[1] == n)
  }
  
  #pick assigned seat if available, else pick random
  choice = ifelse(n %in% seats, n, sample(seats, 1, replace = FALSE))
  remaining_seats <- seats[seats != choice]
  
  #call function again for next passenger
  drunkPassenger(n + 1, remaining_seats)
}


seats <- 2:100
runs <- 1000

drunk_passengers_100 <- map(seats, function(x) {
  data.frame(
    num_passengers = x,
    pct_correct_last_seat = mean(map_lgl(1:runs, ~ drunkPassenger(1, 1:x)))
  )
}) |>
  list_rbind() |>
  mutate(runs = runs, num_seats = num_passengers) |>
  select(num_passengers, num_seats, pct_correct_last_seat, runs)


drunk_passengers_100 |>
  ggplot(aes(x = num_passengers, y = pct_correct_last_seat)) +
    geom_line() +
    ylim(0,1)+
    labs(
      title = "Odds Last Person Gets Their Seat",
      x = "Number of Passengers",
      y = ""
    )

```

So we see on a fully booked flight, there's a 50% chance the last guy gets his seat, no matter how many passengers!

[Part 2: Extra Seats]{style="font-size:1.25em; color:gray;"}

Now lets consider that the plane isn't fully booked, that there's n passengers but k seats, given that k \>= n.

Lets take n to be 20, and simulate k from 20:30

```{r}
drunkPassenger_k <- function(n, passengers, seats){
  #check if last guy gets his seat
  if (length(passengers) == n){
    return(n %in% seats)
  }
  
  #fist seat selection
  if (n == 1) {
    first_choice = sample(seats, 1, replace = FALSE)
    #stop simulation if chooses last persons seat
    if (first_choice == length(passengers)) {
      return(FALSE)
    }
    remaining_seats <-seats[seats != first_choice]
  
  } else{
      #pick assigned seat if available, else pick random
      choice = ifelse(n %in% seats, n, sample(seats, 1, replace = FALSE))
      if (choice == length(passengers)) {
        return(FALSE)
      }
      remaining_seats <-seats[seats != choice]
  }
  
  #call function again for next passenger
  drunkPassenger_k(n + 1, passengers, remaining_seats)
}

runs = 1000
passengers = 20
seats = 20:30


drunk_passengers_k20 <- map(seats, function(x) {
  data.frame(
    num_seats = x,
    pct_correct_last_seat = mean(map_lgl(1:runs, ~ drunkPassenger_k(1, 1:passengers, 1:x)))
  )
}) |>
  list_rbind() |>
  mutate(runs = runs, num_passengers = passengers) |>
  select(num_passengers, num_seats, pct_correct_last_seat, runs)

```

Now lets take n to be 5, and simulate k from 5:15.

```{r}
runs = 1000
passengers = 5
seats = 5:15


drunk_passengers_k5 <- map(seats, function(x) {
  data.frame(
    num_seats = x,
    pct_correct_last_seat = mean(map_lgl(1:runs, ~ drunkPassenger_k(1, 1:passengers, 1:x)))
  )
}) |>
  list_rbind() |>
  mutate(runs = runs, num_passengers = passengers) |>
  select(num_passengers, num_seats, pct_correct_last_seat, runs)

```

```{r}

drunk_k <- bind_rows(drunk_passengers_k5, drunk_passengers_k20)
  
drunk_k |>
  group_by(num_passengers) |>
  mutate(diff = num_seats - num_passengers) |>
  ggplot(aes(x = diff, y = pct_correct_last_seat, color = factor(num_passengers), group = num_passengers)) +
    geom_line() +
    labs(
      color = "Number of Passengers",
      x = "Extra Seats",
      title = "Odds Last Person Gets Their Seat",
      y = "")
```

If n is fixed, increasing k makes the probability of last guy finding his seat approaches 1. Pretty intuitive. It's cool to see that the number of passengers and seats doesn't affect the odds, its the difference between them. n=5, k=8 is the same odds as n=20, k=23, since there's 3 extra seats in both examples. This shows a unique solution that isn't very intuitive.
